---
title: "Data Cleaning In R"
author: "Handerson Coq"
date: '2022-04-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Performing some data cleaning tasks

- Install some packages for simplifying the process

```{r}
install.packages("here")
library(here) # makes referencing files easier
```
```{r}
install.packages("skimr")
library(skimr) # allows you to skim through data
```
```{r}
install.packages("janitor")
library(janitor) # contains functions for cleaning data
```
```{r}
install.packages("dplyr")
library(dplyr) # dplyr contains many functions for data cleaning
```

- Install some data to work with

```{r}
install.packages("palmerpenguins")
```

- Getting a comprehensive summary about the data using the skim_without_charts(), glimpse() and head() functions

```{r}
skim_without_charts(penguins)
glimpse(penguins)
head(penguins)
```

- Make use of pipes and select() to work with specific columns from the data

```{r}
penguins %>% 
  select(species) # select just the species column
```
```{r}
penguins %>% 
  select(-species) # select all columns but the species column
```

- Renaming columns with the rename() function

```{r}
penguins %>% 
  rename(bill_length = bill_length_mm) %>% 
  rename(bill_depth = bill_depth_mm) %>% 
  rename(body_mass = body_mass_g) %>% 
  rename (flipper_length = flipper_length_mm)
```

- Make columns format consistent with the rename_with() function

```{r}
penguins %>% 
  rename_with(toupper) %>%  # make column names all caps
  rename_with(tolower) # make column names all lower cases
```

- make sure that column names are well formatted with the clean_names() function

```{r}
clean_names(penguins)
```

- Combining data in different columns using the unite() function

```{r}
# date_year <- bookings_df %>%
# select(arrival_date_year, arrival_date_month) %>% 
# unite(arrival_month_year, c("arrival_date_month", "arrival_date_year"), sep = " ")
```


